### 2022-12-07 doc/manuscript/translocation.qmd
Update methods and references

<!--# Exact author order to be determined) -->

@@

@@

During the period `r min(d1$year)`-`r max(d1$year)`, we conducted `r n_distinct(d1$site_id, d1$date)` translocations of adult frogs to `r n_distinct(d1$site_id)` recipient sites in Yosemite (`r transno_min`-`r transno_max` translocations per site; @fig-yosemap) <!--# Add table of all translocations: site_id, date, frog #s, donor site, etc.? -->. Each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs collected from one of `r n_distinct(d1$donor)` donor populations (@fig-yosemap). <!--# #fig-yosemap will likely be moved to supplement -->Estimates of frog survival during the one year following translocation indicate that survival was highly variable between sites, but relatively constant within sites (when multiple translocations were made to individual sites; @fig-translocation-survival). This suggests an important effect of site characteristics on frog survival. <!--# Add analysis results: translocation_survival_analysis, analysis 2 -->

The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation (bdload) was not among them (@fig-survival-postdens). This result is consistent with the exome capture results suggesting that frogs in recovering populations have evolved increased resistance and/or tolerance to Bd infection. Important predictors are shown in @fig-cond-effects, and included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996). Males had higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects). In addition to Bd load, effects of frog size (length_z), winter severity in the year of translocation (snow_t), day of year on which translocation was conducted (day), and translocation order (order_first) were all unimportant (@fig-survival-postdens).

@@

Given observed yearly adult survival probabilities, our stage-structured model predicted that six of our twelve translocated populations should experience a long-run growth rate $\lambda$ near or greater than 1 in the presence of Bd (@fig-viability A; median predicted $\lambda$ ranges from 0.95-1.16 for these six populations). When the probability of successful recruitment $\omega$ was greater than 0.1 (the conservative estimate we used as a baseline value in our model), long-run growth rate of these populations could be notably greater than 1 (@fig-viability A). In contrast, for the other six translocated populations where yearly adult survival was less than 0.5, our model predicted that it was unlikely that long-run growth rate in these populations was greater than 1, even when the probability of successful recruitment to the adult life stage was near 0.4 (@fig-viability A). A sensitivity analysis of long-run growth rate $\lambda$ to recruitment probability $\omega$ and yearly adult survival probability of naturally recruited individuals $\sigma_{A_R}$ showed that $\lambda$ was most sensitive to small changes in $\sigma_{A_R}$ (@fig-viability-supp A).

Given demographic stochasticity (i.e., probabilistic births and deaths of individuals in a population) and environmental variability in the probability of adult recruitment $\omega$, our model predicted that, following a a single introduction of 40 adult individuals into a population, the six populations with the highest adult survival probabilities had 50-year extinction probabilities of less than 1 (@fig-viability B) -- in other words, even under conservative assumptions about recruitment probability $\omega$, these populations had some probability of persistence over 50 years following a single introduction. For these six populations, as $\omega$ increased to 0.3, 50 year extinction probability was less 0.5, indicating strong potential for long-term persistence in the presence of Bd and environmental variability in recruitment success (@fig-viability B). In contrast, for the six populations where $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be 1. To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest surveillance history (@fig-viability-supp B; population 70550, surveyed for 16 years).

@@

![Map showing the locations of Bd-naive or recovering frog populations from which DNA samples were collected for the exome capture analyses. Black polygons indicate the boundaries of Yosemite, Kings Canyon, and Sequoia national parks (from top to bottom). To obscure the exact locations of populations, random noise was added to all point coordinates. Inset map shows the location of the study area within California. In both maps, elevation is indicated by the colored hillshade layer (dark green = lowest elevation, white = highest elevation)](figures/map_exome_points.png){#fig-exomemap}

![Map showing the locations of translocated and donor populations in Yosemite National Park (park border indicated by black polygon). To obscure the exact locations of populations, random noise was added to all point coordinates. Elevations are indicated as in @fig-exomemap. See inset map in @fig-exomemap to locate Yosemite National Park within California.](figures/map_translocation_points.png){#fig-yosemap}

![Median survival of each cohort of translocated frogs to the 12 recipient sites included in this study. Sites are arranged along the x-axis using the average by site of the median survival per translocation. Frog survival for the one-year period following each translocation was estimated using capture-mark-recapture methods. Dot colors indicate the donor population from which frogs in each translocated cohort were collected.](figures/translocation_survival_bysiteid.png){#fig-translocation-survival}

![For the best model identifying predictors of 1-year frog survival following translocations, estimated posterior density curves and shaded 95% uncertainty intervals for all predictor variables. Variables are considered important predictors if the associated uncertainty interval does not overlap zero (dashed red line).](figures/mcmc_areas_m1d.png){#fig-survival-postdens}

![Conditional effects of the important predictors of 1-year frog survival: (A) elevation, (B) winter severity in the year following the translocation, (C) sex, and (D) donor population. In (A) and (B), blue lines are medians and gray ribbons are 95% uncertainty intervals. In (C) and (D), box plots show medians, first and third quartiles, largest and smallest values within 1.5x interquartile range, and values outside the 1.5x interquartile range.](figures/cond_effects_plot.png){#fig-cond-effects}

![A. The predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and probability or recruitment $\omega$, given the parameterized, deterministic model for translocated *R. sierrae*. Color points show the predicted $\lambda$ values for the twelve translocated populations when probability of recruitment to adults is $\omega = 0.1$ (given by the dashed line). B. The predicted fifty-year extinction probabilities of the twelve translocated populations given demographic stochasticity, environmental variability in $\omega$ and different mean values of $\omega$. The dashed line shows where $\omega = 0.1$, for comparison to A.](figures/pop_viability_figures_for_manuscript_v1.jpg){#fig-viability}

<!-- Use figure v1 or v2? If v1, sort legend labels by sigma-AR-->

![A. The elasticity of $\lambda$ with changes in three parameters: $\sigma_{A_R}$, $\omega$, and *F*. Elasticity is calculated at the default parameter values for each population and $\omega$ = 0.1. B. The 100 simulated trajectories (black lines) that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (colored line).](figures/pop_viability_figures_for_supp.jpg){#fig-viability-supp}

<!-- In legend for (B), change "lake" to "site". Switch color of 70550 in A so that color of line in B can be easily referenced in text, e.g., "orange". Also, use colorblind-friendly palette for figure, if possible. -->

@@

We used exome capture methods to compare genomes of frogs in populations with different histories of Bd exposure, in an effort to identify putative genes under selection that are linked to reduced frog susceptibility to Bd-caused lethal and sub-lethal effects. Specifically, we compared genomes of frogs that have never experienced a Bd-caused epizootic ("Bd-naive") versus populations that experienced a Bd epizootic during the past several decades and have since recovered to varying degrees ("recovering"). These populations persist in an enzootic state, characterized by high Bd prevalence and Bd infection intensities (load) in adults that are well below the level expected to cause frog mortality. Bd-exposure histories of these populations are based on a decade or more of data from VES and Bd surveillance using skin swabbing.

@@

DNA samples were collected via buccal swabbing of xx-xx frogs from seven Bd-naive populations and seven recovering populations. These populations represented a wide latitudinal range in the southern Sierra Nevada, from northern Yosemite National Park to south of Sequoia National Park (@fig-exomemap). The latitudinal range of the two population types was similar.

@@

Translocations of *R. sierrae* were conducted in Yosemite National Park (California, USA) as part of an effort to reestablish populations of this endangered species. We identified donor populations from which adult frogs ($\geq$ 40 mm snout-vent length) could be collected using several years of visual encounter surveys (VES) and skin swab collections [@knapp2016]. Results indicated that these populations contained several hundred *R. sierrae* adults and hundreds or thousands of tadpoles, and that frog abundances were recovering from previous Bd-caused declines. Bd prevalence in donor populations was high and Bd infection intensity (load) was below the level at which frog mortality is predicted <!--Calculate median prevalence and load -->. Recipient sites to which frogs were translocated were chosen based on previous *R. sierrae* presence (determined from VES and/or museum records) or characteristics that suggested high quality habitat for this species [@knapp2005].

Between 2006 and 2020 we conducted 1-4 translocations of adult frogs to 12 recipient sites (@fig-yosemap). In each translocation, adult frogs ($\geq$ 40 mm snout-vent length) were collected from one of three donor populations and measured, weighed, swabbed, and PIT tagged. Frogs were transported to the recipient site either on foot or via helicopter. Following release, translocated populations were visited one to six times per summer (calculate min, max, mean) and assessed using diurnal capture-mark-recapture (CMR) surveys and VES. CMR surveys allowed quantification of adult frog survival, adult population size, and recruitment of new adults, and VES provided estimates of tadpole and subadult abundance.

@@

During 2006-2012, CMR surveys were conducted on a single day (primary period) per site visit, during which all habitats were searched repeatedly for adult frogs. Frog were captured using handheld nets, identified via their PIT tag (or tagged if they were untagged), measured, weighed, swabbed, and released at the capture location. During 2013-2022, we generally used a robust design in which all habitats were searched during several consecutive days (secondary periods - calculate min, max, mean), and frogs were processed as described above. However, when the number of frogs detected on the first survey day was low, we conducted only a single-day CMR survey. Within a primary period, frogs that were captured during more than one secondary period were measured, weighed, and swabbed only during the first capture, and identified only during subsequent captures. Frog skin swabs were analyzed using standard Bd DNA extraction and qPCR methods. For details of survey and swab analysis methods, see @joseph2018.

During each site visit, we conducted VES either immediately before CMR surveys or during the first day of CMR surveys. VES was conducted by walking the entire water body perimeter, first 100 m of each inlet and outlet stream, and any fringing ponds and wetlands, and counting all *R. sierrae* tadpoles and subadults. These *R. sierrae* life stages have high detectablility, and counts are highly repeatable and provide estimates of relative abundance <!-- reference -->.

@@

To identify important predictors of frog survival following translocation, we developed multilevel models in a Bayesian framework. We performed all analyses using R Statistical Software [v4.4.4; @rsoftware2022] and the rstanarm package [@rstanarm2022].

@@

**Model description**

To determine the implications of acquired Bd resistance on long-term population viability, we developed a dynamic, population model for *R. sierrae*. Our central question was: how does the magnitude and variation in observed adult survival probability across translocated populations affect the long-term persistence probability of populations? We developed a model that tracked seven state variables of a *R. sierrae* population: the density of translocated adults ($A_T$), the density of adults naturally recruited into the population ($A_R$), the density of first year larvae ($L_1$), the density of second year larvae ($L_2$), the density of third year larvae ($L_3$), the density of first year juveniles ($J_1$), and the density of second year juveniles ($J_2$). We divided adults into two classes $A_T$ and $A_R$ because there was evidence that the survival probability of translocated adults and naturally recruited adults differed.

@@

The parameters in this model are yearly survival probability $\sigma_{*}$, where the "\*" indicates a particular state variable, the probability that a female frog reproduces in a given year $p_F$, the number of eggs produced by a female frog in a year that successfully hatch $F$, the probability of a first year tadpole remaining as a tadpole $p_{L1}$, the probability of a second year tadpole remaining as a tadpole $p_{L2}$, and the probability of a first year juvenile remaining as a juvenile $p_{J1}$. The parameter $\omega$ gives the probability of a juvenile successfully recruiting to an adult.

In this model we ignore density-dependent recruitment because we were interested in the growth of the population from an initial reintroduction and whether this growth was sufficient to prevent extinction over 50 years following the introduction. We also did not directly consider the dynamics of Bd in this model. We made this decision because i) all translocated populations are infected with Bd at high prevalence ii) host density does not seem to play a significant role in multi-year Bd infection dynamics in this system (Wilber et al. 2022). Thus, ignoring Bd infection dynamics and instead assuming all host vital rates are in the presence of high Bd prevalence significantly simplifies the model without much loss of realism.

**Model analysis**

**Incorporating yearly variability in vital rates**

We extracted yearly survival probabilities for translocated adults $\sigma_{A_T}$ and naturally recruited adults $\sigma_{A_R}$ from the mark-recapture model described elsewhere. While we observed yearly variability in adult survival *within* a population, the magnitude of this variability was small compared to *among* population variability. Thus, we did not include yearly within-population variability in adult survival in this analysis.

However, within a population there was substantial yearly variability in the successful recruitment of adults, greater than what we would expect from Poisson variability about a mean value. Therefore, we allowed for yearly variability in the probability of successful recruitment $\omega$ (more details below).

**Estimating model parameters**

The baseline parameter values for the model and how they were estimated are given in Table 1. Parameters $\sigma_{A_T}$ and $\sigma_{A_R}$ were extracted directly from our mark-recapture model described elsewhere. For populations where we had a sufficient number of pit-tagged, naturally recruited adults, we observed that $\sigma_{A_T}$ and $\sigma_{A_R}$ could be notably different, with $\sigma_{A_R} > \sigma_{A_T}$. For populations where we had not pit-tagged any naturally recruited adults, we could not directly estimate $\sigma_{A_R}$ and set $\sigma_{A_R} = \sigma_{A_T}$.

**Model analysis and simulation**

We performed four analyses on our model. First, we considered a deterministic version of our model with no yearly heterogeneity in probability of successful recruitment $\omega$ and calculated the predicted long-run growth rate $\lambda$ of a population for different values of $\sigma_{A_R}$ and $\omega$. We then fixed $\omega = 0.1$ and calculated the predicted growth rate of our six populations. Second, we performed a local sensitivity analysis on $\lambda$ with respect to parameters $\omega$, $\sigma_{A_R}$, and $F$ to determine how small changes in these parameters could influence the long-run deterministic growth rate of populations.

Third, we defined version of the model with demographic and environmental stochasticity, where environmental stochasticity was represented by among year variability in $\omega$. We used this model to simulate a one time introduction of 40 translocated adult frogs. We ran this simulation 1000 times for each population and computed the probability of a population becoming extinct after 50 years given the observed parameter values and environmental stochasticity in $\omega$. We varied the mean recruitment probability $\omega$ from 0.05 and 0.3 and drew values of $\omega$ each year from a beta distribution with a dispersion parameter of $\phi = 2$ (when $\omega = 0.5$ and $\phi = 2$ the beta distribution is uniform between 0 and 1).

Finally, we assessed whether our stochastic model could reproduce observed trajectories of population recovery. We focused on population 70550 because this was our longest mark-recapture time series for a translocated population and because this population shows evidence of recovery. We simulated our model for 16 years, repeating the simulation 50,000 times. For each run and each year, we drew $\omega$ from a uniform distribution between 0 and 1 (or a beta distribution with mean 0.5 and $\phi = 2$). Using Approximate Bayesian Computing and rejection sampling (@kosmala2016), we identified the top 2% of trajectories (i.e., 1000 trajectories) that minimized the sum of squared errors between the observed and predicted data. The $\omega$ values associated with these "best" trajectories represented an approximate posterior distribution [@beaumont2010]. Using these best fit trajectories, we assessed whether our model could qualitatively describe the patterns of recovery in the observed data for population 70550.

>>>>

<!--# Exact author order to be determined -->

@@

#### Frog translocations result in population reestablishment despite ongoing Bd infection

@@

During the period `r min(d1$year)`-`r max(d1$year)`, we conducted `r n_distinct(d1$site_id, d1$date)` translocations of adult frogs to `r n_distinct(d1$site_id)` recipient sites in Yosemite (`r transno_min`-`r transno_max` translocations per site; @fig-yosemap) <!--# Add table of all translocations: site_id, date, frog #s, donor site, etc.? -->. Each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs collected from one of `r n_distinct(d1$donor)` donor populations (@fig-yosemap). <!--# #fig-yosemap will likely be moved to supplement -->Estimates of frog survival during the one year following translocation indicate that survival was highly variable between sites, but relatively constant within sites (for the subset of sites that received multiple translocations; @fig-translocation-survival). This suggests an important effect of site characteristics on frog survival. <!--# Add analysis results: translocation_survival_analysis, analysis 2 -->

The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (bd_load; @fig-survival-postdens). This result is consistent with the exome capture results suggesting that frogs in recovering populations have evolved increased resistance and/or tolerance to Bd infection. Important predictors are shown in @fig-cond-effects, and included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996). Males had higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects). In addition to Bd load, effects of frog size (length), winter severity in the year of translocation (snow_t), day of year on which translocation was conducted (day), and translocation order (order_first) were all unimportant (@fig-survival-postdens).

@@

#### Many reestablished populations are likely to persist over the long term

Given observed yearly adult survival probabilities, our stage-structured model predicted that six of twelve translocated populations should experience a long-run growth rate $\lambda$ near or greater than 1 in the presence of Bd (@fig-viability A; median predicted $\lambda$ ranges from 0.95-1.16 for these six populations). When the probability of successful recruitment $\omega$ was greater than 0.1 (the conservative estimate we used as a baseline value in our model), long-run growth rate of these populations could be notably greater than 1 (@fig-viability A). In contrast, for the other six translocated populations where yearly adult survival was less than 0.5, our model predicted that it was unlikely that long-run growth rate in these populations was greater than 1, even when the probability of successful recruitment to the adult life stage was near 0.4 (@fig-viability A). A sensitivity analysis of long-run growth rate $\lambda$ to recruitment probability $\omega$ and yearly adult survival probability of naturally recruited individuals $\sigma_{A_R}$ showed that $\lambda$ was most sensitive to small changes in $\sigma_{A_R}$ (@fig-viability-supp A).

Given demographic stochasticity (i.e., probabilistic births and deaths of individuals in a population) and environmental variability in the probability of adult recruitment $\omega$, our model predicted that, following a a single introduction of 40 adult individuals into a population, the six populations with the highest adult survival probabilities had 50-year extinction probabilities of less than 1 (@fig-viability B). Therefore, even under conservative assumptions about recruitment probability $\omega$, these populations had some probability of persistence over 50 years following a single introduction. For these six populations, as $\omega$ increased from 0.1 to 0.3, 50 year extinction probability was less 0.5, indicating strong potential for long-term persistence in the presence of Bd and environmental variability in recruitment success (@fig-viability B). In contrast, for the six populations where $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be 1. To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest surveillance history (@fig-viability-supp B; population 70550, surveyed for 16 years). <!--# Add information about natural recruitment probability in the study populations? Can we even calculate recruitment probability? -->

@@

![Map showing the locations of Bd-naive or recovering frog populations from which DNA samples were collected for the exome capture analyses. Black polygons indicate the boundaries of Yosemite, Kings Canyon, and Sequoia National Parks (from top to bottom). To obscure the exact locations of populations, random noise was added to all point coordinates. Inset map shows the location of the study area within California. In both maps, elevation is indicated by the colored hillshade layer (dark green = lowest elevation, white = highest elevation)](figures/map_exome_points.png){#fig-exomemap}

![Map showing the locations of translocated and donor populations in Yosemite National Park (park boundary indicated by black polygon). To obscure the exact locations of populations, random noise was added to all point coordinates. Elevations are indicated as in @fig-exomemap. See inset map in @fig-exomemap to locate Yosemite National Park within California. <!--# Add park boundaries to inset map? -->](figures/map_translocation_points.png){#fig-yosemap}

![Median survival of each cohort of translocated frogs to the 12 recipient sites included in this study. Sites are arranged along the x-axis using the average of the median survival per translocation at each site. Frog survival for the one-year period following each translocation was estimated using capture-mark-recapture methods. Dot colors indicate the donor population from which frogs in each translocated cohort were collected.](figures/translocation_survival_bysiteid.png){#fig-translocation-survival}

![For the best model identifying predictors of 1-year frog survival following translocation, estimated posterior density curves and shaded 95% uncertainty intervals for all predictor variables. Variables are considered important predictors if the associated uncertainty interval does not overlap zero (dashed red line).](figures/mcmc_areas_m1d.png){#fig-survival-postdens}

![Conditional effects of the important predictors of 1-year frog survival: (A) elevation of recipient site, (B) winter severity in the year following translocation, (C) sex, and (D) donor population. In (A) and (B), blue lines are medians and gray ribbons are 95% uncertainty intervals. In (C) and (D), box plots show medians, first and third quartiles, largest and smallest values within 1.5x interquartile range, and values outside the 1.5x interquartile range.](figures/cond_effects_plot.png){#fig-cond-effects}

![A. The predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and probability of recruitment $\omega$, given the parameterized, deterministic model for translocated *R. sierrae*. Colored points show the predicted $\lambda$ values for the twelve translocated populations when probability of recruitment into the adult class is $\omega = 0.1$ (given by the dashed line). B. The predicted fifty-year extinction probabilities of the twelve translocated populations given demographic stochasticity, environmental variability in $\omega$, and different mean values of $\omega$. The dashed line shows where $\omega = 0.1$, for comparison to A.](figures/pop_viability_figures_for_manuscript_v1.jpg){#fig-viability}

<!-- Use figure v1 or v2 (omega = 0.3)? If v1, sort legend labels by sigma-AR-->

![A. The elasticity of $\lambda$ with changes in three parameters: $\sigma_{A_R}$, $\omega$, and *F (*number of eggs produced by a female frog in a year that successfully hatch). Elasticity is calculated at the default parameter values for each population and $\omega$ = 0.1. B. The 100 simulated trajectories (black lines) that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (colored line).](figures/pop_viability_figures_for_supp.jpg){#fig-viability-supp}

<!-- In legend for (A), change "lake" to "site". Switch color of 70550 in A so that color of associated line in B can be easily referenced in text, e.g., "orange". Also, use colorblind-friendly palette for figure, if possible. -->

@@

To identify putative genes under selection that are linked to reduced frog susceptibility to Bd-caused lethal and sub-lethal effects, we used exome capture methods to compare genomes of frogs in populations with different histories of Bd exposure. Specifically, we compared genomes of frogs from populations that have never experienced a Bd-caused epizootic ("Bd-naive") versus frogs from populations that experienced a Bd epizootic during the past several decades and have since recovered to varying degrees ["recovering"; @knapp2016, @vredenburg2010dynamics]. These latter populations persist in an enzootic state [@briggs2010], characterized by high Bd prevalence and Bd infection intensities (load) in adults that are typically well below the level expected to cause frog mortality [@knapp2011, @vredenburg2010dynamics]. Bd-exposure histories of these populations are based on a decade or more of data from VES and Bd surveillance using skin swabbing [e.g., @knapp2016].

@@

DNA samples were collected via buccal swabbing [@broquet2007] of xx-xx frogs from seven Bd-naive populations and seven recovering populations. These populations represented a wide latitudinal range in the southern Sierra Nevada, from northern Yosemite National Park to south of Sequoia National Park (@fig-exomemap). To minimize potential biases associated with different latitudinal ranges of the two population types, we selected populations to represent similar latitudinal ranges.

@@

##### Field methods

Translocations of *R. sierrae* were conducted in Yosemite National Park (California, USA) as part of an effort to reestablish populations of this endangered species. We identified donor populations from which adult frogs ($\geq$ 40 mm snout-vent length) could be collected using several years of visual encounter surveys (VES) and skin swab collections [@knapp2016]. Results indicated that these populations contained several hundred *R. sierrae* adults and hundreds or thousands of tadpoles, and that frog abundances were recovering from previous Bd-caused declines [@knapp2016]. Bd prevalence in donor populations was high and Bd infection intensity (load) was below the level at which frog mortality is predicted <!--Calculate median prevalence and load -->. Recipient sites to which frogs were translocated were chosen based on previous *R. sierrae* presence (determined from VES and/or museum records) or characteristics that suggested high quality habitat for this species [@knapp2005].

Between 2006 and 2020, we conducted 1-4 translocations of adult frogs to 12 recipient sites (@fig-yosemap). In each translocation, adult frogs were collected from one of three donor populations and measured, weighed, swabbed, and PIT tagged. Frogs were transported to the recipient site either on foot or via helicopter. Following release, translocated populations were visited one to six times per summer <!--# calculate min, max, mean --> and assessed using diurnal capture-mark-recapture (CMR) surveys and VES. CMR surveys allowed quantification of adult survival, adult population size, and recruitment of new adults, and VES provided estimates of tadpole and subadult abundance.

@@

During 2006-2012, CMR surveys were conducted on a single day (primary period) per site visit, during which all habitats were searched repeatedly for adult frogs. Frog were captured using handheld nets, identified via their PIT tag (or tagged if they were untagged), measured, weighed, swabbed, and released at the capture location. During 2013-2022, we generally used a robust design in which all habitats were searched during several consecutive days (secondary periods <!--# calculate min, max, mean-->), and frogs were processed as described above. However, when the number of frogs detected on the first survey day was low, we conducted only a single-day CMR survey. Within a primary period, frogs that were captured during more than one secondary period were measured, weighed, and swabbed only during the first capture, and during subsequent captures were only identified and released. Frog skin swabs were analyzed using standard Bd DNA extraction and qPCR methods. For details of survey and swab analysis methods, see @joseph2018.

During each site visit, we conducted VES either immediately before CMR surveys or during the first day of CMR surveys. VES was conducted by walking the entire water body perimeter, first 100 m of each inlet and outlet stream, and any fringing ponds and wetlands, and counting all *R. sierrae* tadpoles and subadults. These *R. sierrae* life stages have high detectablility, and counts are highly repeatable and provide estimates of relative abundance [@knapp2000].

##### Estimation of adult survival and population size

@@

##### Predictors of post-translocation frog survival

To identify important predictors of frog survival following translocation, we developed multilevel models in a Bayesian framework. We performed all analyses using R Statistical Software [v4.4.4; @rsoftware2022] and the rstanarm package [@rstanarm2022]. <!--# Add details and citations, from PeerJ paper? -->

@@

##### Model description

To determine the implications of acquired Bd resistance on long-term population viability, we developed a dynamic, population model for *R. sierrae*. Our central question was: how does the magnitude and variation in observed adult survival probability across translocated populations affect the long-term persistence probability of populations? We developed a model that tracked seven state variables of an *R. sierrae* population: the density of translocated adults ($A_T$), the density of adults naturally recruited into the population ($A_R$), the density of first year larvae ($L_1$), the density of second year larvae ($L_2$), the density of third year larvae ($L_3$), the density of first year juveniles ($J_1$), and the density of second year juveniles ($J_2$). We divided adults into two classes $A_T$ and $A_R$ because there was evidence that the survival probability of translocated adults and naturally recruited adults differed.

@@

The parameters in this model are yearly survival probability $\sigma_{*}$, where the "\*" indicates a particular state variable<!--# "*" subscript is not translating correctly in .docx. Shows as upsidedown question mark -->, the probability that a female frog reproduces in a given year $p_F$, the number of eggs produced by a female frog in a year that successfully hatch $F$, the probability of a first year tadpole remaining as a tadpole $p_{L1}$, the probability of a second year tadpole remaining as a tadpole $p_{L2}$, and the probability of a first year juvenile remaining as a juvenile $p_{J1}$. The parameter $\omega$ gives the probability of a juvenile successfully recruiting to an adult.

In this model we ignore density-dependent recruitment because we were interested in the growth of the population from an initial reintroduction and whether this growth was sufficient to prevent extinction over 50 years following the introduction. We also did not directly consider the dynamics of Bd in this model. We made this decision because i) all translocated populations are infected with Bd at high prevalence, and ii) host density does not seem to play a significant role in multi-year Bd infection dynamics in this system [@wilber2022]. Thus, ignoring Bd infection dynamics and instead assuming all host vital rates are in the presence of high Bd prevalence significantly simplifies the model without much loss of realism.

##### Model analysis

###### Incorporating yearly variability in vital rates

We extracted yearly survival probabilities for translocated adults $\sigma_{A_T}$ and naturally recruited adults $\sigma_{A_R}$ from the CMR model used to estimate population-level parameters of translocated populations (see "Estimation of adult survival and population size" above). Although we observed yearly variability in adult survival *within* a population, the magnitude of this variability was small compared to *among* population variability. Thus, we did not include yearly within-population variability in adult survival in this analysis.

However, within a population there was substantial yearly variability in the successful recruitment of adults, greater than what we would expect from Poisson variability about a mean value. Therefore, we allowed for yearly variability in the probability of successful recruitment $\omega$ (additional details provided in "Estimating model parameters" below).

###### Estimating model parameters

The baseline parameter values for the model and how they were estimated are given in Table 1 <!--# Create table --> . Parameters $\sigma_{A_T}$ and $\sigma_{A_R}$ were extracted directly from our CMR model described elsewhere. For populations where we had a sufficient number of pit-tagged, naturally recruited adults, we observed that $\sigma_{A_T}$ and $\sigma_{A_R}$ could be notably different, with $\sigma_{A_R} > \sigma_{A_T}$. For populations lacking any naturally-recruited adults, we were unable to directly estimate $\sigma_{A_R}$, and instead set $\sigma_{A_R} = \sigma_{A_T}$.

###### Model analysis and simulation

We performed four analyses on our model. First, we considered a deterministic version of our model with no yearly heterogeneity in probability of successful recruitment $\omega$, and calculated the predicted long-run growth rate $\lambda$ of a population for different values of $\sigma_{A_R}$ and $\omega$. We then fixed $\omega = 0.1$ and calculated the predicted growth rate of our six populations.

Second, we performed a local sensitivity analysis <!--# referred to elsewhere as "elasticity" -->on $\lambda$ with respect to parameters $\omega$, $\sigma_{A_R}$, and $F$ to determine how small changes in these parameters could influence the long-run deterministic growth rate of populations.

Third, we defined a version of the model with demographic and environmental stochasticity, where environmental stochasticity was represented by among-year variability in $\omega$. We used this model to simulate a one-time introduction of 40 translocated adult frogs. We ran this simulation 1000 times for each population and computed the probability of a population becoming extinct after 50 years given the observed parameter values and environmental stochasticity in $\omega$. We varied the mean recruitment probability $\omega$ from 0.05 and 0.3 and drew values of $\omega$ each year from a beta distribution with a dispersion parameter of $\phi = 2$ (when $\omega = 0.5$ and $\phi = 2$ the beta distribution is uniform between 0 and 1).

Finally, we assessed whether our stochastic model could reproduce observed trajectories of population recovery. We focused on population 70550 because this was our longest CMR time series for a translocated population and because this population shows evidence of substantial post-translocation increases in adult abundance associated with population establishment and recovery. We simulated our model for 16 years, repeating the simulation 50,000 times. For each run and each year, we drew $\omega$ from a uniform distribution between 0 and 1 (or a beta distribution with mean 0.5 and $\phi = 2$). Using Approximate Bayesian Computing and rejection sampling (@kosmala2016), we identified the top 2% of trajectories (i.e., 1000 trajectories) that minimized the sum of squared errors between the observed and predicted data. The $\omega$ values associated with these "best" trajectories represented an approximate posterior distribution [@beaumont2010]. Using these best fit trajectories, we assessed whether our model could qualitatively describe the patterns of recovery in the observed data for population 70550.
